Accessing the nucleon tensor structure in inclusive deep inelastic scattering

### 2016-07-26 1607 - JMC for g2/Paper/JMC_dis.tex
Fixed some other references

The tensor charge is a fundamental property of the nucleon, at
present poorly constrained. It has been estimated in lattice
QCD (see, \eg,
\cite{Green:2012ej,Bali:2014nma,Bhattacharya:2015wna,Abdel-Rehim:2015ow}), but
only limited information is available from direct measurements. The
way to extract the tensor charge from experimental measurements requires first
of all the extraction the so-called transversity parton distribution function,
denoted by $h_1^q(x)$~\cite{Radici:2015mwa,Anselmino:2015sxa,Kang:2015msa}. 
The integral of the transversity distribution
corresponds to the contribution of flavor $q$ to the tensor charge.

@@

The transversity distribution is notoriously difficult to access because it is
a chiral-odd function and needs to be combined with a spin-flip mechanism to
appear in a scattering process~\cite{Jaffe:1996zw,Barone:2001sp}. Usually, this spin flip is provided by another
nonperturbative distribution or fragmentation function, implying that
transversity cannot be accessed in inclusive DIS,
but only in more complex processes such as semi-inclusive DIS or Drell-Yan~\cite{Ralston:1979ys,Jaffe:1991kp,Jaffe:1993xb,Collins:1993kk}.

@@

The correlator $\Xi$ can be parametrized in terms of jet parton correlation
functions (PCFs), using the vectors $l$ and $n_+$: 
\begin{equation}
\Xi(l,n_+) = \Lambda A_1(l^2)\,{\bm 1} + A_2(l^2)\,\lslash 
+ \frac{\Lambda^2}{l \cdot n_{+}} \nslash_+ \, B_{1}(l^2)
+ \frac{i \Lambda}{2 l \cdot n_{+}} [\,\lslash,\nslash_+ ] \, B_{2}(l^2) \ .
\label{e:jetexpansion}
\end{equation} 
Time reversal invariance in QCD requires $B_{2}=0$, while $B_{1}$ contributes
only at twist-4 order, and will not be considered further in this paper. We
focus, instead, on the role of chiral odd terms in the $g_2$ structure function up to twist 3. At this order, 
\begin{equation}
  \Xi(l,n_+) = \Lambda A_1(l^2)\,{\bm 1} + A_2(l^2)\,\lslash 
    + O(\Lambda^2/Q^2)
\label{e:jetexpansion-tw3}
\end{equation} 
is nothing else than the full quark propagator; note however, that we consider
here the full QCD vacuum rather than the perturbative one.  
The $A_1$ and $A_2$ terms can be nicely interpreted in terms of the spectral representation of the cut quark propagator \cite{D'Hoker-lecture-notes,Romao-lecture-notes},
\begin{align} 
  \Xi(l) =  
  \int d \mu^2 \big[ J_1 (\mu^2)\,\mu + J_2 (\mu^2)\,\lslash \big] \,
  \delta(l^2 -\mu_j^2) \ ,
\label{e:jetspectral}
\end{align}
where $\mu^2$ is interpreted as the invariant mass of the current jet, {\it
  i.e.}, of the particles going through the cut in the top blob of
Fig.\ref{fig:handbags}(a), and the $J_i$ are the spectral functions of the
quark propagator, that have been also called ``jet functions'' in
\cite{Accardi:2008ne,SCET?} or ``jet factors''~\cite{Collins:2007ph}. These satisfy
\begin{align}
  J_2(\mu^2) \geq J_1(\mu^2) \geq 0
  \hspace*{0.5cm} \text{and} \hspace*{0.5cm}
  \int d\mu^2 J_2(\mu^2) = 1 \ .
\label{eq:jetfnsprops}
\end{align}
From a comparison of Eqns.\eqref{e:jetexpansion} and \eqref{e:jetspectral}, one can see that 
\begin{align}
  A_1(l^2)&=\frac{\sqrt{l^2}}{\Lambda}J_1(l^2) & A_2(l^2)&=J_2(l^2) \ .
  \label{eq:jet_vs_spectral}
\end{align}

@@

When inserting the jet correlator in the handbag diagram for inclusive DIS, the invariant jet mass $\mu^2$ is integrated from 0 to $Q^2(1/x_B-1)$. This induces (kinematical) corrections of order $O(1/Q^2)$, whose effect on the $F_2$ structure function has been studied in Ref.~\cite{Accardi:2008ne}:

@@

It is important to notice that $\xi_2=1$ exactly due to CPT invariance \cite{Weinberg-book}, while $0 < \mj < \int d\mu^2 \mu J_2(\mu^2)$ is dynamically determined. From the analytic properties of spectral functions we may expect \cite{Accardi:2008ne} $J_2(\mu^2) = Z \delta(\mu^2-m_q) + \bar J_2 (\mu^2) \theta (\mu^2-m_\pi^2)$ with the continuum starting at $m_\pi$, the mass of the pion, due to color confinement effects. Taking into account that $J_1 < J_2$, we may therefore expect 

@@

% As we shall see, an analogous formula for quark-gluon-quark correlators is not needed.
%The $\xi_1$ term is chiral-odd and therefore can appear in the inclusive cross section only coupled to the transversity function $h_1$. Therefore, for our analysis 
The relevant part of the semi-inclusive hadronic tensor for our analysis is 
\todo{[AA] I am using the notation in Piet Mulder's lecture notes - this will need to be checked.}
\begin{align}
  \label{eq:Wsidis_ini}
  2 \Lambda  W^{\mu\nu}
    & = i \frac{2\Lambda}{Q} \hat t^{[\mu}_{\phantom \perp} 
    \epsilon_\perp^{\nu]\rho}S_{\perp\rho} 
    \sum_q e_q^2
    \bigg[ 2 x_b g_T(x_B) \sum_h \int dz d^2p_{hT} D_1^{q,h}(z,p_{hT}) 
  %\\ &
  + 2 h_1(x_B) \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) \bigg] + \ldots
\end{align}
Here and we reintroduced the quark flavor $q$ for maximum clarity, and $e_q$ its electric charge.
The first term can be easily integrated with the help of the sum rule \eqref{eq:SIDIS_to_DIS_TMDlevel}. To integrate the latter, we first need make use of the relation $\tilde E(z) = E(z) - (\mq/\Lambda) z D_1(z)$, which is a consequence of the QCD equations of motion \cite{Bacchetta:2006tn}, then utilize the sum rule \eqref{eq:SIDIS_to_DIS}:
\begin{align}
  \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) 
    = \sum_h \int dz d^2p_{hT} \Big[ E^{q,h}(z,p_{hT}) - \frac{\mq}{\Lambda} z D_1^{q,h}(z,p_{hT}) \Big]
    = \xi_1 - \frac{\mq}{\Lambda} \xi_2 = \frac{\mj - \mq}{\Lambda} \ .
\end{align}
This formula is the single most important result of this paper, and provides a
non perturbative generalization of the commonly used $\int\tilde E =0$ sum
rule introduced in \cite{Jaffe:1996zw}. Indeed, calculating the jet correlator 
on the perturbative vacuum one would obtain, as already discussed, $\mj=\mq$
and the new term would vanish.

@@

The new term in Eq.\eqref{e:FLTint} only appears in the $g_2$ structure function. Following the derivation in Ref.~\cite{ABMS}, one finds
\begin{align}
\label{e:g2}
  g_2(x_B) = g_2^{WW} + \frac{1}{2}\,\sum_a e_a^2
\biggl(
    \widetilde g_T^{a \star}(x) 
    + \int_x^1\frac{dy}{y} \widehat{g}_T^q(y) 
    + \frac{\mq}{\Lambda} \left(\frac{h_1^q}{x}\right)^\star(x) 
    + \frac{\mj-\mq}{\Lambda} \frac{h_1^q(x)}{x} 
\Biggr) \ ,
\end{align}
where we defined $f^*(x) = -f(x) + \int_x^1\frac{dy}{y} f(y)$. The first 4
terms coincide with the result obtained in the conventional handbag
approximation \cite{ABMS}, while the fifth is new.

@@

It is important to estimate the size of teh variou scontributions to the non Wandzura-Wilczek part of $g_2$. We define the shorthand notation
\begin{align}
g_2^{tw3} & = \frac{1}{2}\,\sum_a e_a^2
    \biggl(
    \widetilde g_T^{a \star}(x) 
    + \int_x^1\frac{dy}{y} \widehat{g}_T^q(y) 
    \biggr) 
&
g_2^{\text{quark}} &= \frac{1}{2}\,\sum_a e_a^2 
 \frac{\mq}{\Lambda} (h_1^q/x)^\star(x),
&
g_2^{\text{jet}} &= \frac{1}{2}\,\sum_a e_a^2 
\frac{\mj-\mq}{\Lambda} \frac{h_1^q(x)}{x}. 
\end{align} 
These are comapred in Figure~\ref{f:g2contrib} to the $g_2-g_2^{WW}$ function
obtained in the very recent JAM15 fit of polarized DIS asymmetries
\cite{Sato:2016tuz}, that includes a large amount of precise data at large $x$
from Jefferson Lab, and simultaneously fits the higher-twist components of in
$g_1$ and $g_2$ to the data. For the ``pure twist-3'' contribution,
$g_2^{tw3}$, {\it i.e.}, the contribution from quark-gluon-quark matrix
elements, we show a model calculation by Braun et al. \cite{Braun-et-al}; for
other estimates, see \cite{Other-tw3-calcs}. To estimate the contributions
from quark ($g_2^q$) and jet mass ($g_2^{jet}$) effects, that depend on chiral
odd quark-quark matrix elements, we use the recent Pavia15 fit of the
transversity distribution from Ref.~\cite{Radici:2015mwa}, which is comparable
also to other 
extractions~\cite{Anselmino:2013vqa,Kang:2015msa}. Furthermore, we choose the
values of the mass parameters to be $\mq=5$ MeV and $\mj = 100$ MeV.

@@

In the neutron case, the jet contribution is very negative at intermediate to large values of $x$. If one trusts the order of magnituse of the $g_2^{tw3}$ calculation by Braun et al., one would conclude that the jet contribution cannot be that large. However, the latteris dominated by the $d$ quark's transversity, whose fit suffers from large systematic uncertainties and saturates the negative Soffer bound.Recent deata in $p+p$ collisions indicate, however, that $h_1^d$ is less negative than in the Pavia15 fits, in agreement with the JAM15 fit of the non Wandzura-Wilczek contribution to $g_2$. Correspondingly the jet contribution to the proton at $x \approx 0.1$ would become less positive, inproving as well teh agreement with the JAM15 fit.

\begin{itemize}
\item smaller-$x$: constrains the small-$x$ behavior of transversity.
\end{itemize}
Some ideas to check:
\begin{itemize}
\item JAM13 neutron has almost 0 twist 3, with small positive contribution at small x. Can we use this to highlight teh $g_2^{jet}$ contribution? Or maybe as a cnstrain: if Braun is right then $g_2^{jet}(n) \approx -g_2^{tw3}(n)$. (still, does not help much for the proton).
\item In Jam 13, Fig. 6, it looks like in general the violation of the BC sum rule is a small-x effect --> constraints on h(x) fits? This cannot really be checked vs. JAM15, that assumes $g_2^(0)$ respects BC. 
\end{itemize}

\begin{figure}[tbh]
\begin{center}
\includegraphics[width=8cm]{g2contrib}
\includegraphics[width=8cm]{g2contribN}
\caption{\label{f:g2contrib} 
Different contributions to the non Wandzura-Wilczek part of the proton (left)
and neutron (right) $g_2$ structure function compared to the JAM15 fit of the
$g_2-g_2^{\text{WW}}$ (solid black) \cite{Sato:2016tuz}. The quark and jet
contributions are shown with a dotted red and a dot-dashed green line
respectively, with uncertainty bands coming form the Pavia15 fit of the
transversity function \cite{Radici:2015mwa}. The unceratainty in the choice
$m_q=5$ GeV and $M_q=100$ GeV is not shown. The pure twist-3 contribution
calculated by Braun et al. \cite{Braun:2011aw} is shown as a dashed blue line
(no uncertainty estiamte was provided in teh original reference). 
}
\end{center}
\end{figure}

@@

The zero-th moment corresponds to an extension of teh Burkhardt-Cottingham sum rule \cite{Burkhardt:1970ti,Jaffe:1996zw}, that appears to be broken by the non-perturbative spin-flip contribution from the jet function: 
\begin{align}
  \label{eq:BC}
  \int dx\, g_2(x) = \frac{\mj-\mq}{\Lambda} \int dx\, \frac{1}{x} h_1(x) \ .
\end{align}
This sum rule, in which the pure twist-3 part does not take part, is a consequence of the Lorentz invariance of the DIS cross section, that entails $\int_0^1 dx g_1^a(x) = \int_0^1 g_T^q(x)$. It explicitly displays a contribution from spin-flip processes, that are included in the original derivation by Burkhardt and Cottingham \cite{Burkhardt:1970ti} but do not show up in treatments that only consider free field quark propagators for the struck quark \cite{Jaffe:1996zw}.

@@

The first moment is the first in which a contribution from the pure twist-3 part of $g_2$ appears:
\begin{align}
  d_1 & = \frac12 \sum_q e_q^2 \bigg( 2 \tilde g_T^q[1] + \hat g_T^q[1]
    + \frac{2M_q-m_q}{\Lambda} h_1^q[1] \bigg)
\end{align}
where $h_1^q[1] = \int_0^1 \dx $ is the contribution of a quark $q$ to the tensor charge. In this case, the transevrsity contribution is a background to the extraction of teh poure twist-3 piece. Fortunately, it is a quantity that can be extracted from the lattice \cite{Bhattacharya:2016zcn} or fitted \cite{Pavia15, others, reviews?}. Furthermore the new sum rule \eqref{eq:ABsumrule} and bound \eqref{eq:ABbound} promise to improve future transversity fits. What is less obvious is how to calculate $M_q-m_q$ from first principles, since this is related to the spectral function of the quark propagator. We will, however, discuss later perspectives on how to measure it. Therefore the pure twist-3 part can, in principle, be properly isolated.

The third moment is interesting because the pure twist-3 part can be related to quark-gluon-quark correlators, see \cite{Jaffe-lectures}, and interpreted as as the average color force experienced by the struck quark as it exits teh nucleon \cite{M.Burkardt}. As in the previous case, however, the transversity contribution needs to be subtracted:
\begin{align}
  d_2 & = \frac12 \sum_q e_q^2 \bigg( 3 \tilde g_T^q[2] + \hat g_T^q[2]
    + \frac{3M_q-m_q}{\Lambda} h_1^q[2] \bigg)
\end{align}

>>>>

The tensor charge is a fundamental property of the nucleon, at
present poorly constrained. It has been estimated in lattice
QCD (see, \eg,
\cite{Green:2012ej,Bali:2014nma,Bhattacharya:2015wna,Abdel-Rehim:2015owa,Bhattacharya:2016zcn}), but
only limited information is available from direct measurements. The
way to extract the tensor charge from experimental measurements requires first
of all the extraction the so-called transversity parton distribution function,
denoted by $h_1^q(x)$ (see Ref.~\cite{Barone:2001sp} for a review on transversity and
Refs.~\cite{Radici:2015mwa,Anselmino:2015sxa,Kang:2015msa} for the most recent
extractions). 
The integral of the transversity distribution
corresponds to the contribution of flavor $q$ to the tensor charge.
The knowledge of the tensor charge can be used also to put constraints on the search of physics beyond the Standard
Model~\cite{Cirigliano:2013xha,Bhattacharya:2015esa,Courtoy:2015haa}.

@@

The transversity distribution is notoriously difficult to access because it is
a chiral-odd function and needs to be combined with a spin-flip mechanism to
appear in a scattering process~\cite{Jaffe:1996zw}. Usually, this spin flip is provided by another
nonperturbative distribution or fragmentation function, implying that
transversity cannot be accessed in inclusive DIS,
but only in more complex processes such as semi-inclusive DIS or Drell-Yan~\cite{Ralston:1979ys,Jaffe:1991kp,Jaffe:1993xb,Collins:1993kk}.

@@

The correlator $\Xi$ can be parametrized in terms of jet parton correlation
functions (PCFs), using the vectors $l$ and $n_+$: 
\begin{equation}
\Xi(l,n_+) = \Lambda A_1(l^2)\,{\bm 1} + A_2(l^2)\,\lslash 
+ \frac{\Lambda^2}{l \cdot n_{+}} \nslash_+ \, B_{1}(l^2)
+ \frac{i \Lambda}{2 l \cdot n_{+}} [\,\lslash,\nslash_+ ] \, B_{2}(l^2) \ .
\label{e:jetexpansion}
\end{equation} 
Time reversal invariance in QCD requires $B_{2}=0$, while $B_{1}$ contributes
only at twist-4 order, and will not be considered further in this paper. We
focus, instead, on the role of chiral odd terms in the $g_2$ structure function up to twist 3. At this order, 
\begin{equation}
  \Xi(l,n_+) = \Lambda A_1(l^2)\,{\bm 1} + A_2(l^2)\,\lslash 
    + O(\Lambda^2/Q^2)
\label{e:jetexpansion-tw3}
\end{equation} 
is nothing else than the full quark propagator; note however, that we consider
here the full QCD vacuum rather than the perturbative one.  
The $A_1$ and $A_2$ terms can be nicely interpreted in terms of the spectral representation of the cut quark propagator \cite{D'Hoker-lecture-notes,Romao-lecture-notes},
\begin{align} 
  \Xi(l) =  
  \int d \mu^2 \big[ J_1 (\mu^2)\,\mu + J_2 (\mu^2)\,\lslash \big] \,
  \delta(l^2 -\mu_j^2) \ ,
\label{e:jetspectral}
\end{align}
where $\mu^2$ is interpreted as the invariant mass of the current jet, {\it
  i.e.}, of the particles going through the cut in the top blob of
Fig.\ref{fig:handbags}(a), and the $J_i$ are the spectral functions of the
quark propagator, that have been also called ``jet functions'' in
\cite{Accardi:2008ne,SCET?}. These satisfy
\begin{align}
  J_2(\mu^2) \geq J_1(\mu^2) \geq 0
  \hspace*{0.5cm} \text{and} \hspace*{0.5cm}
  \int d\mu^2 J_2(\mu^2) = 1 \ .
\label{eq:jetfnsprops}
\end{align}
From a comparison of Eqns.\eqref{e:jetexpansion} and \eqref{e:jetspectral}, one can see that 
\begin{align}
  A_1(l^2)&=\frac{\sqrt{l^2}}{\Lambda}J_1(l^2) & A_2(l^2)&=J_2(l^2) \ .
  \label{eq:jet_vs_spectral}
\end{align}

@@

When inserting the jet correlator in the handbag diagram for inclusive DIS,
the invariant jet mass $\mu^2$ is integrated from 0 to $Q^2(1/x_B-1)$. This
induces (kinematical) corrections of order $O(1/Q^2)$, whose effect on the
$F_2$ structure function has been studied in Ref.~\cite{Accardi:2008ne}: 

@@

It is important to notice that $\xi_2=1$ exactly due to CPT invariance
\cite{Weinberg:1995mt}, while $0 < \mj < \int d\mu^2 \mu J_2(\mu^2)$ is
dynamically determined. From the analytic properties of spectral functions we
may expect \cite{Accardi:2008ne} $J_2(\mu^2) = Z \delta(\mu^2-m_q) + \bar J_2
(\mu^2) \theta (\mu^2-m_\pi^2)$ with the continuum starting at $m_\pi$, the
mass of the pion, due to color confinement effects. Taking into account that
$J_1 < J_2$, we may therefore expect  

@@

% As we shall see, an analogous formula for quark-gluon-quark correlators is not needed.
%The $\xi_1$ term is chiral-odd and therefore can appear in the inclusive cross section only coupled to the transversity function $h_1$. Therefore, for our analysis 
The relevant part of the semi-inclusive hadronic tensor for our analysis is 
\begin{align}
  \label{eq:Wsidis_ini}
  2 \Lambda  W^{\mu\nu}
    & = i \frac{2\Lambda}{Q} \hat t^{[\mu}_{\phantom \perp} 
    \epsilon_\perp^{\nu]\rho}S_{\perp\rho} 
    \sum_q e_q^2
    \bigg[ 2 x_b g_T(x_B) \sum_h \int dz d^2p_{hT} D_1^{q,h}(z,p_{hT}) 
  %\\ &
  + 2 h_1(x_B) \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) \bigg] + \ldots
\end{align}
Here and we reintroduced the quark flavor $q$ for maximum clarity, and $e_q$ its electric charge.
The first term can be easily integrated with the help of the sum rule
\eqref{eq:SIDIS_to_DIS_TMDlevel}. To integrate the latter, we first need make
use of the relation $\tilde E(z) = E(z) - (\mq/\Lambda) z D_1(z)$, which is a
consequence of the QCD equations of motion \cite{Bacchetta:2006tn}, then
utilize the sum rule \eqref{eq:SIDIS_to_DIS}: 
\begin{align}
  \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) 
    = \sum_h \int dz d^2p_{hT} \Big[ E^{q,h}(z,p_{hT}) - \frac{\mq}{\Lambda} z D_1^{q,h}(z,p_{hT}) \Big]
    = \xi_1 - \frac{\mq}{\Lambda} \xi_2 = \frac{\mj - \mq}{\Lambda} \ .
\end{align}
This formula is the single most important result of this paper, and provides a
non perturbative generalization of the commonly used $\int\tilde E =0$ sum
rule introduced in \cite{Jaffe:1996zw}. Indeed, calculating the jet correlator 
on the perturbative vacuum one would obtain, as already discussed, $\mj=\mq$
and the new term would vanish.

@@

The new term in Eq.\eqref{e:FLTint} only appears in the $g_2$ structure function. Following the derivation in Ref.~\cite{Accardi:2009au}, one finds
\begin{align}
\label{e:g2}
  g_2(x_B) = g_2^{WW} + \frac{1}{2}\,\sum_a e_a^2
\biggl(
    \widetilde g_T^{a \star}(x) 
    + \int_x^1\frac{dy}{y} \widehat{g}_T^q(y) 
    + \frac{\mq}{\Lambda} \left(\frac{h_1^q}{x}\right)^\star(x) 
    + \frac{\mj-\mq}{\Lambda} \frac{h_1^q(x)}{x} 
\Biggr) \ ,
\end{align}
where we defined $f^*(x) = -f(x) + \int_x^1\frac{dy}{y} f(y)$. The first 4
terms coincide with the result obtained in the conventional handbag
approximation \cite{Accardi:2009au}, while the fifth is new.

@@

It is important to estimate the size of the variou scontributions to the non Wandzura-Wilczek part of $g_2$. We define the shorthand notation
\begin{align}
g_2^{tw3} & = \frac{1}{2}\,\sum_a e_a^2
    \biggl(
    \widetilde g_T^{a \star}(x) 
    + \int_x^1\frac{dy}{y} \widehat{g}_T^q(y) 
    \biggr) 
&
g_2^{\text{quark}} &= \frac{1}{2}\,\sum_a e_a^2 
 \frac{\mq}{\Lambda} (h_1^q/x)^\star(x),
&
g_2^{\text{jet}} &= \frac{1}{2}\,\sum_a e_a^2 
\frac{\mj-\mq}{\Lambda} \frac{h_1^q(x)}{x}. 
\end{align} 
These are comapred in Figure~\ref{f:g2contrib} to the $g_2-g_2^{WW}$ function
obtained in the very recent JAM15 fit of polarized DIS asymmetries
\cite{Sato:2016tuz}, that includes a large amount of precise data at large $x$
from Jefferson Lab, and simultaneously fits the higher-twist components of in
$g_1$ and $g_2$ to the data. For the ``pure twist-3'' contribution,
$g_2^{tw3}$, {\it i.e.}, the contribution from quark-gluon-quark matrix
elements, we show a model calculation by Braun et al.~\cite{Braun:2011aw}; for
other estimates, see \cite{Other-tw3-calcs}. To estimate the contributions
from quark ($g_2^q$) and jet mass ($g_2^{jet}$) effects, that depend on chiral
odd quark-quark matrix elements, we use the recent Pavia15 fit of the
transversity distribution from Ref.~\cite{Radici:2015mwa}, which is comparable
also to other 
extractions~\cite{Anselmino:2013vqa,Kang:2015msa}. Furthermore, we choose the
values of the mass parameters to be $\mq=5$ MeV and $\mj = 100$ MeV.

@@

In the neutron case, the jet contribution is very negative at intermediate to
large values of $x$. If one trusts the order of magnituse of the $g_2^{tw3}$
calculation by Braun et al., one would conclude that the jet contribution
cannot be that large. However, the latteris dominated by the $d$ quark's
transversity, whose fit suffers from large systematic uncertainties and
saturates the negative Soffer bound.Recent deata in $p+p$ collisions indicate,
however, that $h_1^d$ is less negative than in the Pavia15 fits, in agreement
with the JAM15 fit of the non Wandzura-Wilczek contribution to
$g_2$. Correspondingly the jet contribution to the proton at $x \approx 0.1$
would become less positive, inproving as well the agreement with the JAM15
fit.

\begin{itemize}
\item smaller-$x$: constrains the small-$x$ behavior of transversity.
\end{itemize}
Some ideas to check:
\begin{itemize}
\item JAM13 neutron has almost 0 twist 3, with small positive contribution at
  small x. Can we use this to highlight the $g_2^{jet}$ contribution? Or maybe
  as a cnstrain: if Braun is right then $g_2^{jet}(n) \approx
  -g_2^{tw3}(n)$. (still, does not help much for the proton). 
\item In Jam 13, Fig. 6, it looks like in general the violation of the BC sum
  rule is a small-x effect --> constraints on h(x) fits? This cannot really be
  checked vs. JAM15, that assumes $g_2^(0)$ respects BC.  
\end{itemize}

\begin{figure}[tbh]
\begin{center}
\includegraphics[width=8cm]{g2contrib}
\includegraphics[width=8cm]{g2contribN}
\caption{\label{f:g2contrib} 
Different contributions to the non Wandzura-Wilczek part of the proton (left)
and neutron (right) $g_2$ structure function compared to the JAM15 fit of the
$g_2-g_2^{\text{WW}}$ (solid black) \cite{Sato:2016tuz}. The quark and jet
contributions are shown with a dotted red and a dot-dashed green line
respectively, with uncertainty bands coming form the Pavia15 fit of the
transversity function \cite{Radici:2015mwa}. The unceratainty in the choice
$m_q=5$ GeV and $M_q=100$ GeV is not shown. The pure twist-3 contribution
calculated by Braun et al. \cite{Braun:2011aw} is shown as a dashed blue line
(no uncertainty estiamte was provided in the original reference). 
}
\end{center}
\end{figure}

@@

The zero-th moment corresponds to an extension of the Burkhardt-Cottingham sum
rule \cite{Burkhardt:1970ti,Jaffe:1996zw}, that appears to be broken by the
non-perturbative spin-flip contribution from the jet function:  
\begin{align}
  \label{eq:BC}
  \int dx\, g_2(x) = \frac{\mj-\mq}{\Lambda} \int dx\, \frac{1}{x} h_1(x) \ .
\end{align}
This sum rule, in which the pure twist-3 part does not take part, is a
consequence of the Lorentz invariance of the DIS cross section, that entails
$\int_0^1 dx g_1^a(x) = \int_0^1 g_T^q(x)$. It explicitly displays a
contribution from spin-flip processes, that are included in the original
derivation by Burkhardt and Cottingham \cite{Burkhardt:1970ti} but do not show
up in treatments that only consider free field quark propagators for the
struck quark \cite{Jaffe:1996zw}.

@@

The first moment is the first in which a contribution from the pure twist-3 part of $g_2$ appears:
\begin{align}
  d_1 & = \frac12 \sum_q e_q^2 \bigg( 2 \tilde g_T^q[1] + \hat g_T^q[1]
    + \frac{2M_q-m_q}{\Lambda} h_1^q[1] \bigg)
\end{align}
where $h_1^q[1] = \int_0^1 dx $ is the contribution of a quark $q$ to the
tensor charge. In this case, the transevrsity contribution is a background to
the extraction of the poure twist-3 piece. Fortunately, it is a quantity that
can be extracted from the lattice \cite{Green:2012ej,Bali:2014nma,Bhattacharya:2015wna,Abdel-Rehim:2015owa,Bhattacharya:2016zcn} or fitted
\cite{Radici:2015mwa,Anselmino:2015sxa,Kang:2015msa}. Furthermore the new sum rule
\eqref{eq:ABsumrule} and bound \eqref{eq:ABbound} promise to improve future
transversity fits. What is less obvious is how to calculate $M_q-m_q$ from
first principles, since this is related to the spectral function of the quark
propagator. We will, however, discuss later perspectives on how to measure
it. Therefore the pure twist-3 part can, in principle, be properly isolated.

The third moment is interesting because the pure twist-3 part can be related
to quark-gluon-quark correlators, see \cite{Jaffe:1996zw}, and interpreted as
as the average color force experienced by the struck quark as it exits the
nucleon~\cite{Burkardt:2012sd}.  As in the previous case, however, the
transversity contribution needs to be subtracted: 
\begin{align}
  d_2 & = \frac12 \sum_q e_q^2 \bigg( 3 \tilde g_T^q[2] + \hat g_T^q[2]
    + \frac{3M_q-m_q}{\Lambda} h_1^q[2] \bigg)
\end{align}
